---
title: "BiofilmQuality"
author: "Melissa DeSiervo"
date: "5/28/2019"
output: html_document
---

This is the R code for assessing food quality of biofilms for the Greenland 2018 dataset. 

The data are from several raw files

1= CN ratio of biofilm from lab analysis
2= The fatty acid biomass from biofilm samples (from PLFA analysis)
3= Fatty acids categorized by biomarkers (as determined by literature)

And one derived file (from the Pond metrics RMD)


First, upload the data
```{r Upload data}

#1

CNratiodetritus <-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/2_Inputs/CNdetritus.csv?token=AL4COXLWQUTBGUQNPPADDNS634FXW"))

CNratiodetritus <-as.data.frame(CNratiodetritus )

#2 
Fattyacids <-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/2_Inputs/Fattyacids_raw.csv?token=AL4COXLUD6VM74U3LVOS5CC634FV2"))

Fattyacids <-as.data.frame(Fattyacids)


#3 
Biomarkers <-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/2_Inputs/Fattyacids_biomarker.csv?token=AL4COXL7ZAQKUZ2QAHBZRR2634FUA"))

Biomarkers <-as.data.frame(Biomarkers)


#4 
Foodexperiment <-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/2_Inputs/Foodexperiment_2018.csv?token=AL4COXNBEKMWH2TJ24XNCAK634FZ2"))

Foodexperiment <-as.data.frame(Foodexperiment)

#5
pondmetrics <-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/4_Outputs/abioticbypond.csv?token=AL4COXJW4AEO3GXCTITGG2S634FOQ"))

pondmetrics <-as.data.frame(pondmetrics)

```


packages
```{r Upload packages}

library(dplyr)
library(MuMIn)
library(ggplot2)
library(plyr)
library(tidyr)

```



ANOVA tests size, % N
```{r Calculate by pond}

hist(CNratiodetritus$Npercent)

Npercentm1<-lm(Npercent~Pond, data=CNratiodetritus)


Foodexperiment1<-subset(Foodexperiment, Treatment=="DET1")

hist(Foodexperiment1$Winglenth)

Winglengthtm1<-lm(Winglenth~Pond, data=Foodexperiment1)
dayspupm2<-lm(dayspup~Pond, data=Foodexperiment1)


```





Averages and standard error by pond
```{r Calculate by pond}

CNratiobypond<-CNratiodetritus%>% group_by(Pond, Pond2) %>% dplyr::summarise(meanN=mean(Nmg,na.rm=TRUE),seN=(sd(Nmg,na.rm=TRUE)/sqrt(n())),meanNper=mean(Npercent,na.rm=TRUE),seNper=(sd(Npercent,na.rm=TRUE)/sqrt(n())), meanCper=mean(Cpercent,na.rm=TRUE),seCper=(sd(Cpercent,na.rm=TRUE)/sqrt(n())), meanCNratio=mean(Cnratio,na.rm=TRUE),seCNratio=(sd(Cnratio,na.rm=TRUE)/sqrt(n())))%>%arrange(Pond2)

Microbialbiomassbypond<-Fattyacids%>% group_by(Pond, Pond2) %>% dplyr::summarise(Microbialbiomass=mean(Totalbiomass,na.rm=TRUE),seMicrobialbiomass=(sd(Totalbiomass ,na.rm=TRUE)/sqrt(n())))

Foodexperiment1<-subset(Foodexperiment, Treatment=="DET1")

##make new wing lengthF column -- convert M to F and leave F as is###

FoodexperimentMF<-Foodexperiment1%>% mutate(WinglengthF=(ifelse(Sex=="M", (-0.71+(1.12*Winglenth)), Winglenth)))

FEmodel<-lm(WinglengthF~Pond, data=FoodexperimentMF)


Foodexperimentbypond<-FoodexperimentMF%>% group_by(Pond) %>% dplyr::summarise (n = n(), countadult = sum(is.na(daysalive)), meandayspup=mean(dayspup, na.rm=TRUE), sddayspup=sd(dayspup, na.rm=TRUE), sedayspup=(sd(dayspup ,na.rm=TRUE)/sqrt(n())), meanWinglengthF=mean(WinglengthF, na.rm=TRUE), sdWinglengthF=sd(WinglengthF, na.rm=TRUE),seWinglengthF=(sd(WinglengthF ,na.rm=TRUE)/sqrt(n())))%>% mutate(survival=countadult/n)

Foodexperimetbypond<-as.data.frame(Foodexperimentbypond)

Foodexperimetbypond$Pond<-factor(Foodexperimetbypond$Pond, levels=c("East","NoOil","Oil", "Golf", "Waterfall", "Vulgaris", "Vulgaris.small", "Ice"))

```



Fattyacidsbiomarkers

1 go from wide to long
2 merge with the table that has the biomarkers
3 get rid of zeros
4 FATTY ACIDS BY EACH GROUP (18 w 7 is categorized twice)
5 back to wide form
6 autotroph to heterotroph ratio
7 merge back together
8 calculate averages by pond (because I have three samples for golf and vulg)

```{r Merge food quality}

#1
fattyacid_long <- gather(Fattyacids, Fattyacid, Biomass, "12:00":"24:1 w9c", factor_key=TRUE)


#2
fattyacid_biomarkers <- plyr::join(fattyacid_long, Biomarkers, by = "Fattyacid")

#3

fattyacid_biomarkers_2<-filter(fattyacid_biomarkers, Biomass > 0)

#4
fattyacidsbypond_BM1<-fattyacid_biomarkers_2%>% group_by(ID, Pond, Pond2, Biomarker1) %>% dplyr::summarise(TotalBiomass=sum(Biomass))
fattyacidsbypond_BM2<-fattyacid_biomarkers_2%>% group_by(ID, Pond, Pond2, Biomarker2) %>% dplyr::summarise(TotalBiomass=sum(Biomass))

#5 
fattyacidbm1wide<-spread(fattyacidsbypond_BM1, key = Biomarker1, value = TotalBiomass)
fattyacidbm1wide<-plyr::rename(fattyacidbm1wide, replace=c("Algae, cyanobacteria, diatoms" ="PP"))

fattyacidbm2wide<-spread(fattyacidsbypond_BM2, key = Biomarker2, value = TotalBiomass)
fattyacidbm2wide<- plyr::rename(fattyacidbm2wide, replace=c("Algae, cyanobacteria, diatoms" ="PP"))

#6
fattyacidbm1wide<- mutate(fattyacidbm1wide, Ratio1 = PP / Bacteria)
fattyacidbm2wide<- mutate(fattyacidbm2wide, Ratio2 = PP / Bacteria)

#7
fattyacidbm1wide_1<-fattyacidbm1wide%>%select(ID, Pond, PP, Bacteria, Ratio1)
fattyacidbm2wide_2<-fattyacidbm2wide%>%select(ID, Pond, PP, Bacteria, Ratio2)

ratiosbysample <- merge(fattyacidbm1wide_1,fattyacidbm2wide_2,by=c("ID","Pond", "Pond2"))

#8 

ratiosbypond<-ratiosbysample%>% group_by(Pond, Pond2) %>% dplyr::summarise(meanBacteria1=mean(Bacteria.x, na.rm=TRUE), sdBactereia1=sd(Bacteria.x, na.rm=TRUE), meanBacteria2=mean(Bacteria.y, na.rm=TRUE), sdBacterua=sd(Bacteria.y, na.rm=TRUE), meanPP1=mean(PP.x, na.rm=TRUE), sdPP1 = sd(PP.x, na.rm=TRUE), meanPP2=mean(PP.y, na.rm=TRUE), sdPP2 = sd(PP.y, na.rm=TRUE),meanRatio1=mean(Ratio1,na.rm=TRUE), meanRatio2=mean(Ratio2), seratio1=sd(Ratio1, na.rm=TRUE)/sqrt(n()), seratio2 = sd(Ratio2, na.rm=)/sqrt(n()))



```



Merge food quality with pond metrics
```{r Merge food quality}

Foodqualitymetrics<-CNratiobypond %>% right_join(Microbialbiomassbypond, by=c("Pond","Pond2")) 

Foodqualitymetrics_2<-Foodqualitymetrics %>% right_join(ratiosbypond, by=c("Pond","Pond2")) 

Foodqualitymetrics_2$Pond<-plyr::revalue(Foodqualitymetrics_2$Pond, c("Vulgaris small"="Vulgaris.small", "No Oil"="NoOil"))

Foodexperimentbypond$Pond<-plyr::revalue(Foodexperimentbypond$Pond, c("Vulgaris small"="Vulgaris.small", "No Oil"="NoOil"))

Foodqualitymetrics_2<-Foodqualitymetrics_2%>% right_join(Foodexperimentbypond, by=c("Pond")) 

Foodqualitypondmetrics3<-Foodqualitymetrics_2 %>% right_join(pondmetrics, by=c("Pond"))


```


1) Merge and test for correllations between Food quality metrics (5 variables: Percent N, Microbial biomass, Wing length, development days, in food bioassay, auto:het ratio)
2) run a PCA
3) Pull out the PC1 and 2 scores to make a plot
4) Pull out the loadings. 
    PC1 = 66% of the variation. And all 3 variables are about equally related to PC1 in an interpretable way. Higher values on PC1 = higher N, wing lenght in the food bioassay, and microbial biomass in the PLFAs. 
5) Make the PC plot, add on the loadings as vectors


```{r correllation and PCA food quality variables}

#1
Foodqualitymetrics4<-as.data.frame(Foodqualitymetrics_2) %>% select(c(meanWinglengthF,meandayspup, meanNper, Microbialbiomass, meanRatio1))

cor(Foodqualitymetrics4)
pairs(Foodqualitymetrics4)


#2
Foodquality.pca <- prcomp(Foodqualitymetrics4, center = TRUE,scale. = TRUE, retx=TRUE) ##variables are scaled to each other##

summary(Foodquality.pca) ##proportion explained by components##
ev <- Foodquality.pca$sdev^2 ##eigvenvalues##
loadings<-Foodquality.pca$rotation ##loadings##

#3
Foodqualitypcascores<-as.data.frame(Foodquality.pca$x)

Foodqualitypcascores2<-cbind(as.data.frame(Foodqualitymetrics_2), Foodqualitypcascores)
Foodqualitypcascores3<-mutate(Foodqualitypcascores2, PC11=PC1*-1) ####multiplied PC1 by -1 so that its more easily interpretable##

##export as a .csv##


#4
Foodqualitypcaloadings<-as.data.frame(Foodquality.pca$rotation)


```






```{r Make models to predict Food quality PC1}

FoodqualitypondmetricsPCAscores <- merge(Foodqualitypondmetrics3,Foodqualitypcascores3, by = "Pond2")
FoodqualitypondmetricsPCAscores <- FoodqualitypondmetricsPCAscores[, !duplicated(colnames(FoodqualitypondmetricsPCAscores))]

##subset out just the predictors#
subsetpredicators<- subset(FoodqualitypondmetricsPCAscores, select =c(Perim_avg,Area_avg,Area_loss_perc,meanDepth,meanDOC,meanNH4, meanpH,
meanN03N02,meanlogTotalP,meanlogConduc,meanDoPerc,meantemp,meanFPOM_liter))

biofilmdepnames<-subset(FoodqualitypondmetricsPCAscores, select=c("Pond2", "PC11"))
scaledpredictors<-scale(subsetpredicators)
allvariablesscaled<-as.data.frame(cbind(biofilmdepnames, scaledpredictors))

#1
smat3 <- abs(cor(subsetpredicators)) <= .5 #makes a correllation matrix and to say TRUE every time a value is greater than 0.5##
smat3[!lower.tri(smat3)] <- NA ##replaces the duplications with NA#

#2
fullmodel<-lm(PC11~Perim_avg+Area_avg+Area_loss_perc+meanDepth+meanFPOM_liter+meanDOC+meanNH4+meanN03N02+meanlogTotalP+meanlogConduc+meanDoPerc+meanpH+meantemp, na.action = "na.fail", data=FoodqualitypondmetricsPCAscores)

fullmodelscaled<-lm(PC11~Perim_avg+Area_avg+Area_loss_perc+meanDepth+meanFPOM_liter+meanDOC+meanNH4+meanN03N02+meanlogTotalP+meanlogConduc+meanDoPerc+meanpH+meantemp, na.action = "na.fail", data=allvariablesscaled)

#3
dredgedmodels<- dredge(fullmodel, subset=smat3, m.lim=c(1,4))
bestmodelset<-subset(dredgedmodels, delta <2) 

dredgedmodelsscaled<-dredge(fullmodelscaled,subset=smat3, m.lim=c(1,4))
bestmodelsetscaled<-subset(dredgedmodelsscaled, delta<2)

###phosphorus is the best predictor of PC1 (Food quality)

```
