---
title: "Mosquito, biofilm, predator"
author: "Melissa DeSiervo"
date: "8/1/2019"
output: html_document
---

This is the R code for calculating combining mosquito and beetle metrics with biofilm metrics. All of the files used here are derived from previous .rmds and are in the "outputs" folder

First, upload the data
```{r Upload data}

mosquitometrics <- read.csv("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/mosquitometrics.csv")

mosquitometrics2<-dplyr::select(mosquitometrics, c(Site,Pond2, meanTotalliter,mortality, meancumemerg, Winglength, meaneggs, Lambdatotalpop, Nt1, middens))

beetlemetrics <- read.csv("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/beetlemetrics.csv")

beetlemetrics2<-dplyr::select(beetlemetrics, c(Site, Pond2, meanbeetleTotalliter))

biofilmmetrics <- read.csv("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/biofilmmetrics.csv")

biofilmmetrics2<-dplyr::select(biofilmmetrics, c(Site = Pond, Pond2, meanProd.ug.cm.day, meanGP.ug.C.day, controlAFDM))

foodqualitymetrics <- read.csv("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/foodqualitymetrics.csv")

foodqualitymetrics2<-dplyr::select(foodqualitymetrics, c(Pond2, PC11))

abioticbypond<- read.csv("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/abioticbypond.csv")

abioticbypond<-abioticbypond %>% mutate(Pond2=Pond)
abioticbypond$Pond2<- plyr::revalue(abioticbypond$Pond2, c("East"="1", "NoOil"="1a", "Oil"="1b", "Golf"="1c", "Waterfall"="2a", "Vulgaris"="3", "Vulgaris.small"="3a", "Ice"=4))

```

Install necessary packages
```{r Upload packages}
library(dplyr)
library(plyr)
```


```{r Merge them together}

allmetrics<-mosquitometrics2 %>% right_join(beetlemetrics2, by=c("Pond2")) 
allmetrics_2<-allmetrics %>% right_join(biofilmmetrics2, by=c("Pond2")) 

foodqualitymetrics2$Site<-plyr::revalue(foodqualitymetrics2$Site, c("Vulgaris small"="Vulgaris.small"))

allmetrics_3<-allmetrics_2 %>% right_join(foodqualitymetrics2, by=c("Pond2")) 

allmetricsabiotic_3<-allmetrics_3 %>% right_join(abioticbypond, by=c("Pond2")) 


write.csv(allmetrics_3, "C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/populationmetricsaug2019.csv")

write.csv(allmetricsabiotic_3, "C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/populationmetricsabioticaug2019.csv")

```





```{r linear models check assumptions}

par(mfrow = c(2, 2))

densfood<-lm(log(meanTotalliter)~PC11, data=allmetrics_3)
plot(densfood)
summary(densfood)

mortdens<-lm(-1*mortality~meanTotalliter, data=allmetrics_3)
plot(mortdens)
summary(mortdens)

mortprod<-lm(-1*mortality~meanProd.ug.cm.day, data=allmetrics_3)
plot(mortprod)
summary(mortprod)

adultslarvae<-lm(log(meancumemerg)~log(meanTotalliter), data=allmetrics_3)
plot(adultslarvae)
summary(adultslarvae)

mortbeetle<-lm(-1*mortality~meanbeetleTotalliter, data=allmetrics_3)
plot(mortbeetle)
summary(mortbeetle)

sizeprod<-lm(Winglength~meanProd.ug.cm.day, data=allmetrics_3)
plot(sizeprod)
summary(sizeprod)

sizestock<-lm(Winglength~controlAFDM, data=allmetrics_3)
plot(sizestock)
summary(sizestock)

sizequality<-lm(Winglength~PC11, data=allmetrics_3)
plot(sizequality)
summary(sizequality)

sizedensity<-lm(Winglength~log(meanTotalliter), data=allmetrics_3)
plot(sizedensity)
summary(sizedensity)

sizebeetle<-lm(Winglength~meanbeetleTotalliter, data=allmetrics_3)
plot(sizebeetle)
summary(sizebeetle)

grazingdens<-lm(meanGP.ug.C.day~log(middens), data=allmetrics_3)
plot(grazingdens)
summary(grazingdens)

lambdamort<-lm(Lambdatotalpop~mortality, data=allmetrics_3)
plot(lambdamort)
summary(lambdamort)


```




```{r graph significant relationships}


mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=8, colour = "black"))+theme(axis.text.y=element_text(size=8, colour = "black"))+theme(axis.title=element_text(size=10))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.position = "none")

##Figure 5##

Perimeterplot<-ggplot(allmetricsabiotic_3, aes(Perim_avg, meanProd.ug.cm.day)) +geom_point(size=0.8)+geom_text(aes(label=Pond2),position=position_nudge(y=-0.22), size=3)+xlab(bquote(atop("Pond perimeter" (m^{2}))))+ylab(bquote(atop("Biofilm productivity" , ( ~ μg  ~C %.%  ~ cm^{-2} %.% ~day^{-1}))))+ggtitle("")+ stat_smooth(method="lm", se=FALSE, color = "black",size=0.6,)+ mytheme+xlim(30, 300)+ylim(-1, 6)+ annotate("text", x = 37, y = 5.75, label = "a)")


TotalPplot<-ggplot(allmetricsabiotic_3, aes(x=meanlogTotalP, y=PC11)) + geom_point(size=0.8)+geom_text(aes(label=Pond2),position=position_nudge(y=-0.2), size=3)+xlab(bquote(atop("ln(Total P (µg/ L))")))+ylab("Food quality (PC1)")+ggtitle("")+ stat_smooth(method="lm", se=FALSE, color = "black",size=0.6,)+ mytheme+ylim(-4, 3)+xlim(1.25, 3.75)+ annotate("text", x = 1.35, y = 2.7, label = "b)")

jpeg("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/Figure5.jpeg", width = 6, height = 3, units = 'in', res = 600)
grid.arrange(Perimeterplot, TotalPplot, ncol=2, widths = c(1.08, 1))
dev.off()


##Figure 6##


PC11abundance<-ggplot(allmetrics_3, aes(PC11, log(meanTotalliter))) +geom_point(size=0.8)+geom_text(aes(label=Pond2),position=position_nudge(y=-0.2), size=3) + mytheme +xlim(-3.5,3)+ylim(2,5.5)+geom_smooth(color="black",method='lm', size=0.5, se=FALSE)+ylab(bquote(atop("Density of early instars", ln (ind/L), "" )))+xlab(bquote(atop("Food quality (PC1)", )))+ annotate("text", x = -3, y = 5.3, label = "a)")

lnNmortality7<-ggplot(allmetrics_3, aes(log(meanTotalliter), -1*mortality)) +geom_point(size=0.8)+geom_text(aes(label=Pond2),position=position_nudge(y=-0.010), size=3)+xlim(1.75,5.5)+ylim(-0.01, 0.17)+xlab("") + ylab("")+ mytheme +geom_smooth(color="black", method='lm', size=0.5, se=FALSE)+ annotate("text", x = 1.9, y = 0.16, label = "b)")+xlab(bquote(atop("Density of early instars", ln (ind/L))))+ylab(bquote(atop("Larval mortality " , (ind  %.% ~ ind^{-1} %.% ~day^{-1}))))

lnNemergence8<-ggplot(allmetrics_3, aes(log(meanTotalliter), log(meancumemerg))) +geom_point(size=0.8)+geom_text(aes(label=Pond2),position=position_nudge(y=-0.2), size=3)+xlab("")+xlim(1.75,5.5)+ylim(2,6)+ mytheme+ylab("") + annotate("text", x = 1.9, y = 5.9, label = "c)")+geom_smooth(color="black", method='lm', size=0.5,se=FALSE)+xlab(bquote(atop("Density of early instars", ln (ind/L))))+ylab(bquote(atop("Density of emerging adults" , ln (ind/L))))

ga<- ggplotGrob(PC11abundance)
gb <- ggplotGrob(lnNmortality7)
gc <- ggplotGrob(lnNemergence8)

gfig6<-cbind(ga, gb, gc, size = "first")

gfig6abc<-gtable_rbind(ga, gb, gc, size = "max")

jpeg("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/Figure6abclong.jpeg", width = 3, height = 7, units = 'in', res = 800)
grid.arrange(gfig6abc)
dev.off()

##Figure 7##

mortalitylambda<-ggplot(allmetrics_3, aes(-1*mortality, Lambdatotalpop)) +geom_point(size=0.8)+geom_text(aes(label=Pond2),position=position_nudge(y=-0.09), size=3) + mytheme +xlim(-0.02, 0.17)+ylim(0, 2.7) +geom_smooth(color="black",method='lm', size=0.6, se=FALSE)+xlab(bquote(atop("Larval mortality " (ind %.% ~ind^{-1} %.% ~day^{-1 }))))+ylab(bquote(atop("Population growth rate", (lambda))))+ annotate("text", x = -0.015, y = 2.6, label = "a)")

NvsNt<-ggplot(allmetrics_3, aes(meanTotalliter, Nt1)) +geom_point(size=0.8)+geom_text(aes(label=Pond2),position=position_nudge(y=-4), size=3)+xlim(0, 150)+ylim(0, 150)+xlab(bquote(atop("Density of early instars", (ind/L))))+ylab(bquote(atop("Density of early instars next year", (ind/L)))) + mytheme +geom_abline(slope=1, linetype="dashed")+ annotate("text", x = 5, y = 145, label = "b)")

g11<- ggplotGrob(mortalitylambda)
g12 <- ggplotGrob(NvsNt)

gall<-cbind(g11,g12, size = "last")
grid.arrange(gall)

jpeg("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/Figure7ab.jpeg", width = 6, height = 3, units = 'in', res = 600)
grid.arrange(gall)
dev.off()

```




predatormortality<-ggplot(allmetrics_3, aes(meanbeetleTotalliter, -1*mortality)) +geom_point(size=0.8)+geom_text(aes(label=Pond2),position=position_nudge(y=-0.010), size=3)+ylim(-0.01, 0.17)+xlab("") + ylab("")+ mytheme +xlab(bquote(atop("Predator Density", (ind/L))))+ylab(bquote(atop("Larval mortality " , (ind  %.% ~ ind^{-1} %.% ~day^{-1}))))


jpeg("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/predmortality.jpeg", width = 4, height = 3, units = 'in', res = 600)
predatormortality
dev.off()



```{r predator models}

m1<-lm(log(meancumemerg)~mortality, data=allmetrics_3)
m2<-lm(log(meancumemerg)~log(Nfinalliter), data=allmetrics_3)
m3<-lm(log(meancumemerg)~meanbeetleTotalliter, data=allmetrics_3)
m12<-lm(log(meancumemerg)~mortality+log(Nfinalliter), data=allmetrics_3)
m13<-lm(log(meancumemerg)~mortality+meanbeetleTotalliter, data=allmetrics_3)
m23<-lm(log(meancumemerg)~log(Nfinalliter)+meanbeetleTotalliter, data=allmetrics_3)
m123<-lm(log(meancumemerg)~mortality+log(Nfinalliter)+meanbeetleTotalliter, data=allmetrics_3)

##compare  models##
rawaic<-AIC(m1, m2, m3, m12, m13, m23, m123)
nR<-dim(allmetrics_3)[1]  #Sample size 
aictable(rawaic,nR)




```


