---
title: "Pond metrics"
author: "Melissa DeSiervo"
date: "5/1/2019"
output: html_document
---
This is the R code for calculating pond metrics from 2018 Greenland dataset. 

Some data are organized by sampling station within pond (YSI data, FPOM, Soluble nutrients). Other data are organized by pond (Temperature, Perimeter). The original data files are in the "raw data" folder which contains the metadata. 

The data that are uploaded here are from the "inputs" folder. For simplicity, I compiled the YSI data, FPOM, and Soluble nutrients data into ONE .csv called "allpondmetricsbysamplingstation". Temperature and perimeter/area data are loaded in from seperate .csv files

The 3 input data files for this set of analyses are as follows
1) Allpondmetricsbysamplingstation ==(YSI, FPOM and soluble nutrients data) (8 ponds X 3 dates: pond thaw, intermediate instar stage, 50% pupation)
2) Ponds2018temp == Temp data from hobo loggers (8 ponds X hourly data, starting at pond thaw)
3) perimeterareaponds_2018 == perimeter and area data (8 ponds X 2 dates: pond thaw and after emergence)


```{r Upload data and fix date formats}


library (readr)


allpondmetricsbysamplingstation<-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/2_Inputs/allpondmetricsbysamplingstation.csv?token=AL4COXLXLGM66NHYLYB52JK633VTY"))

allpondmetricsbysamplingstation<-as.data.frame(allpondmetricsbysamplingstation)

allpondmetricsbysamplingstation$Date <- as.Date(allpondmetricsbysamplingstation$Date, "%m/%d/%Y")


Ponds2018Temp2<-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/2_Inputs/Allponds2018andairport_Temp.csv?token=AL4COXP6EK25QFTLZ4WR5HC633XRI"))


Ponds2018Temp2<-as.data.frame(Ponds2018Temp2)

Ponds2018Temp2$DateTime<-as.POSIXct(Ponds2018Temp2$DateTime, format = "%m/%d/%Y %H:%M") ##dateformatfordate_time_start##
Ponds2018Temp2$Date <- as.Date(Ponds2018Temp2$Date, "%m/%d/%Y")


perimeterarea<-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/2_Inputs/perimeterandareaponds_2018.csv?token=AL4COXJXOJODNODLCNRBKZS633XYU"))

perimeterarea<-as.data.frame(perimeterarea)

perimeterarea$Pond <- as.factor(perimeterarea$Pond)
perimeterarea$Date <- as.Date(perimeterarea$Date, "%m/%d/%Y")

```

Install necessary packages
```{r Upload packages}

library(dplyr)
library(lme4)
library(chillR)
library(tidyr)
```

For the pond variables where I have sampling station within pond, summarize means and stdev by pond and date 

```{r summarize by pond and date}
abioticsummarybypondanddate<-allpondmetricsbysamplingstation %>% group_by(Pond, Date) %>% dplyr::summarise(meanDepth=mean(Depth, na.rm=TRUE),sdDepth=sd(Depth, na.rm=TRUE),seDepth=sd(Depth, na.rm=TRUE)/sqrt(n()), meanDOC=mean(DOC, na.rm=TRUE),sdDOC=sd(DOC, na.rm=TRUE), seDOC=sd(DOC, na.rm=TRUE)/sqrt(n()), meanNH4=mean(NH4, na.rm=TRUE),sdNH4=sd(NH4, na.rm=TRUE), seNH4=sd(NH4, na.rm=TRUE)/sqrt(n()), meanN03N02=mean(N03N02, na.rm=TRUE),sdN03N02=sd(N03N02, na.rm=TRUE),seN03N02=sd(N03N02, na.rm=TRUE)/sqrt(n()), meanTotalP=mean(TotalP, na.rm=TRUE),sdTotalP=sd(TotalP, na.rm=TRUE),seTotalP=sd(TotalP, na.rm=TRUE)/sqrt(n()),meanConduc=mean(Conduc, na.rm=TRUE),sdConduc=sd(Conduc, na.rm=TRUE),seConduc=sd(Conduc, na.rm=TRUE)/sqrt(n()), meanDoPerc=mean(DoPerc, na.rm=TRUE),sdDoPerc=sd(DoPerc, na.rm=TRUE),seDoPerc=sd(DoPerc, na.rm=TRUE)/sqrt(n()), meanpH=mean(pH, na.rm=TRUE),sdpH=sd(pH, na.rm=TRUE),sepH=sd(pH, na.rm=TRUE)/sqrt(n()), meanFPOM_liter=mean(FPOM_liter, na.rm=TRUE),sdFPOM_liter=sd(FPOM_liter, na.rm=TRUE),seFPOM_liter=sd(FPOM_liter, na.rm=TRUE)/sqrt(n()))

abioticsummarybypond<-allpondmetricsbysamplingstation %>% group_by(Pond) %>% dplyr::summarise(meanDepth=mean(Depth, na.rm=TRUE),sdDepth=sd(Depth, na.rm=TRUE), meanDOC=mean(DOC, na.rm=TRUE),sdDOC=sd(DOC, na.rm=TRUE), meanNH4=mean(NH4, na.rm=TRUE),sdNH4=sd(NH4, na.rm=TRUE), meanN03N02=mean(N03N02, na.rm=TRUE),sdN03N02=sd(N03N02, na.rm=TRUE), meanTotalP=mean(TotalP, na.rm=TRUE),sdTotalP=sd(TotalP, na.rm=TRUE),meanConduc=mean(Conduc, na.rm=TRUE),sdConduc=sd(Conduc, na.rm=TRUE), meanDoPerc=mean(DoPerc, na.rm=TRUE),sdDoPerc=sd(DoPerc, na.rm=TRUE), meanpH=mean(pH, na.rm=TRUE),sdpH=sd(pH, na.rm=TRUE), meanFPOM_liter=mean(FPOM_liter, na.rm=TRUE),sdFPOM_liter=sd(FPOM_liter, na.rm=TRUE))
```


Mixed anova for variables with bonferroni adjusment (Table S2). Subject = Sampling station, Within-subject = Date, Between subject = Pond
    bonferroni adjustment....a0.05=0.0056, a0.01=0.0011, a0.001=0.00011116
```{r anovas bonferroni}

library(car)

allpondmetricsbysamplingstation$SS2<-as.factor(allpondmetricsbysamplingstation$SS2)

Depthmodel<-aov(Depth ~ Pond*Date + Error(SS2), data=allpondmetricsbysamplingstation, na.action=na.omit)

DOCmodel<-aov(DOC ~ Pond*Date + Error(SS2), data=allpondmetricsbysamplingstation, na.action=na.omit)

FPOMmodel<-aov(FPOM_adjust ~ Pond*Date + Error(SS2), data=allpondmetricsbysamplingstation, na.action=na.omit)

NH4model<-aov(NH4 ~ Pond*Date + Error(SS2), data=allpondmetricsbysamplingstation, na.action=na.omit)

N03N02model<-aov(N03N02 ~ Pond*Date + Error(SS2), data=allpondmetricsbysamplingstation, na.action=na.omit)

TotalPmodel<-aov(logTotalP ~ Pond*Date + Error(SS2), data=allpondmetricsbysamplingstation, na.action=na.omit)

pHmodel<-aov(pH ~ Pond*Date + Error(SS2), data=allpondmetricsbysamplingstation, na.action=na.omit)

conductmodel<-aov(logConduc ~ Pond*Date + Error(SS2), data=allpondmetricsbysamplingstation, na.action=na.omit)


Domodel<-aov(DoPerc ~ Pond*Date + Error(SS2), data=allpondmetricsbysamplingstation, na.action=na.omit)


summary(Depthmodel)

summary(FPOMmodel)

summary(DOCmodel)

summary(NH4model)

summary(N03N02model)

summary(TotalPmodel)

summary(pHmodel)

summary(conductmodel)

summary(Domodel)



```

Next, deal with the temperature (Hourly hobo logger data)
1) First turn the data into long format
2) Calculate average temperatures for each day
3) Calculate degree hours with a base of 0, 1, 2, 3, 4, 5
4) Add up each column by date and then divide by 24 to get daily thermal sums
5) Then cumulation by date to get daily thermal sums
6) merge dataframes
7) pull out the relevant thermal sums

```{r calculate thermal sums}
library(tidyr)

#1
Ponds2018Templong<- Ponds2018Temp2%>% gather(Oil, NoOil, East, Golf, Ice, Vulgaris, Vulgaris.small, Waterfall, Airport, key = "Pond", value = "TempC")
Ponds2018Templong$Pond<-as.factor(Ponds2018Templong$Pond)

#2
avgdailytemp<-Ponds2018Templong %>% drop_na()  %>% group_by(Date, Pond) %>%dplyr::summarise(meantemp=mean(TempC,na.rm=TRUE),sdtemp=sd(TempC, na.rm=TRUE)) %>% arrange(Pond, Date)

#3
Degreehoursall2018<-Ponds2018Templong %>% drop_na() %>% mutate(GDD0=ifelse(TempC>0, TempC-0, 0), GDD1=ifelse(TempC>1, TempC-1, 0), GDD2=ifelse(TempC>2, TempC-2, 0), GDD3=ifelse(TempC>3, TempC-3, 0), GDD4=ifelse(TempC>4, TempC-4, 0), GDD5=ifelse(TempC>5, TempC-5, 0))

#4
Degreedaysall2018<-Degreehoursall2018 %>% group_by(Date, Pond) %>%dplyr::summarize(DegreeDay0=(sum(GDD0, na.rm=FALSE))/24, DegreeDay1=(sum(GDD1, na.rm=FALSE))/24, DegreeDay2=(sum(GDD2, na.rm=FALSE))/24, DegreeDay3=(sum(GDD3, na.rm=FALSE))/24, DegreeDay4=(sum(GDD4, na.rm=FALSE))/24, DegreeDay5=(sum(GDD5, na.rm=FALSE))/24) %>% arrange(Pond, Date = as.Date(Date, "%d-%m-%Y"))

#5
Thermalsumsall2018<-Degreedaysall2018  %>% group_by(Pond)%>% arrange(Pond, Date)%>% dplyr::mutate(TS0=cumsum(DegreeDay0),TS1=cumsum(DegreeDay1), TS2=cumsum(DegreeDay2), TS3=cumsum(DegreeDay3), TS4=cumsum(DegreeDay4), TS5=cumsum(DegreeDay5))

#6
tempandthermalsums <- avgdailytemp %>% right_join(Thermalsumsall2018, by=c("Date","Pond"))

#7
Ponddates<-abioticsummarybypondanddate %>% select(Pond, Date)
Ponddates$Pond<-plyr::revalue(Ponddates$Pond, c("Vulgaris small"="Vulgaris.small"))
thermalsumsatdates<-Ponddates %>% left_join(tempandthermalsums, by=c("Date","Pond"))%>% select(Pond, Date, TS0)%>% arrange(Pond, Date)
```


Next, pull out the relevant temperatures 
1) Pull out the avg temp from the first day of sampling
2) Pull out the dates in between sampling dates 1 and 2 and take average
3) Pull out the dates in between sampling dates 2 and 3 and take average
4) Merge the average temps
```{r temp dates}

#1
tempfirstdateeachpond <-  tempandthermalsums%>% group_by(Pond) %>%  slice(which.min(Date))%>%  select(Date, Pond, meantemp)

#2
betweenfirstandsecond <- tempandthermalsums %>% filter(Pond=="East" & Date>="2018-05-18" & Date<="2018-06-01"|
  Pond=="Golf" & Date>="2018-05-19" & Date<="2018-06-02"|Pond=="Ice" & Date>="2018-05-20" & Date<="2018-06-04"|
  Pond=="NoOil" & Date>="2018-05-15" & Date<="2018-05-27"|Pond=="Oil" & Date>="2018-05-09" & Date<="2018-05-13"|
  Pond=="Vulgaris" & Date>="2018-05-17" & Date<="2018-05-29"|Pond=="Vulgaris.small" & Date>="2018-05-20" & Date<="2018-06-04"|
  Pond=="Waterfall" & Date>="2018-05-16" & Date<="2018-05-27")

avgtempbetweenfirstandsecond<-betweenfirstandsecond  %>% group_by(Pond) %>%dplyr::summarize(avgtemp12=mean(meantemp), sdtemp12=sd(meantemp))

#3
betweensecondandthird <- tempandthermalsums %>% filter(Pond=="East" & Date>="2018-06-01" & Date<="2018-06-17"|
  Pond=="Golf" & Date>="2018-06-02" & Date<="2018-06-11"|Pond=="Ice" & Date>="2018-06-04" & Date<="2018-06-12"|
  Pond=="NoOil" & Date>="2018-05-27" & Date<="2018-06-04"|Pond=="Oil" & Date>="2018-05-13" & Date<="2018-06-05"|
  Pond=="Vulgaris" & Date>="2018-05-29" & Date<="2018-06-13"|Pond=="Vulgaris.small" & Date>="2018-06-04" & Date<="2018-06-13"|
  Pond=="Waterfall" & Date>="2018-05-27" & Date<="2018-06-11")

avgtempbetweensecondandthird<-betweensecondandthird  %>% group_by(Pond) %>%dplyr::summarize(avgtemp23=mean(meantemp), sdtemp23=sd(meantemp))

#4
firstandfirstsecond <- tempfirstdateeachpond %>% right_join(avgtempbetweenfirstandsecond, by=c("Pond"))%>% select(Pond, meantemp, avgtemp12)
allavgtemps<-firstandfirstsecond  %>% right_join(avgtempbetweensecondandthird, by=c("Pond"))%>% select(Pond, meantemp, avgtemp12, avgtemp23)

#5
avgtempsatsamplingdates<- allavgtemps%>% gather(meantemp, avgtemp12, avgtemp23, key = "Samplingperiod", value = "AvgTempC")%>% arrange(Pond)
avgtempsatsamplingdates$Samplingperiod<-plyr::revalue(avgtempsatsamplingdates$Samplingperiod, c("meantemp"=1, "avgtemp12"=2, "avgtemp23"=3))

```



Add temperature and thermal sum data to the Pond X date dataframe

```{r merge temp with other abiotic variables}

abioticsummarybypondanddate<-abioticsummarybypondanddate %>% arrange(Pond, Date)
avgtempsatsamplingdates<-avgtempsatsamplingdates %>% arrange(Pond, Samplingperiod)

abioticsummarybypondanddate2<-data.frame(abioticsummarybypondanddate,avgtempsatsamplingdates$AvgTempC )
abioticsummarybypondanddate2 <- abioticsummarybypondanddate2 %>% dplyr::rename(AvgTemp=avgtempsatsamplingdates.AvgTempC)

abioticsummarybypondanddate2$Pond<-plyr::revalue(abioticsummarybypondanddate2$Pond, c("Vulgaris small"="Vulgaris.small"))

abioticsummarybypondanddate3<- abioticsummarybypondanddate2 %>% left_join(thermalsumsatdates, by=c("Pond", "Date"))

```


Reorganize data to make correllations
1)spread to wide format perimeter and area 
2)Calculate average perimeter, area, and loss
3)Calculate total average temp for each pond different than table 1
4)Summarize other abiotic variables by pond (Use the log transformed version for total P and conductivity)
5)Merge into final data table by pond
6)get rid of some columns
```{r correllations at pond level}
#1
perimeterwide<-perimeterarea %>% select(Pond,Measurement,Perimeter) %>% spread(Measurement,Perimeter)%>% rename(Perimeter_1=Beginning, Perimeter_2=End)

areawide<-perimeterarea %>% select(Pond,Measurement,Area) %>% spread(Measurement,Area)%>% rename(Area_1=Beginning, Area_2=End)

perimandareawide<- perimeterwide %>% left_join(areawide, by=c("Pond"))

#2
perimandareawide2<-perimandareawide %>% mutate(Perim_avg=(Perimeter_1 + Perimeter_2)/2, Area_avg=(Area_1+Area_2)/2, Perim_loss=abs(Perimeter_2-Perimeter_1), Area_loss=abs(Area_2-Area_1), Area_loss_perc=(Area_loss/Area_1)*100)

#3
avgtempearlyinstartopupae <- tempandthermalsums %>% filter(Pond=="East" & Date>="2018-05-18" & Date<="2018-06-18"|
  Pond=="Golf" & Date>="2018-05-19" & Date<="2018-06-11"|Pond=="Ice" & Date>="2018-05-20" & Date<="2018-06-12"|
  Pond=="NoOil" & Date>="2018-05-15" & Date<="2018-06-04"|Pond=="Oil" & Date>="2018-05-09" & Date<="2018-06-05"|
  Pond=="Vulgaris" & Date>="2018-05-17" & Date<="2018-06-13"|Pond=="Vulgaris.small" & Date>="2018-05-20" &     Date<="2018-06-13"|Pond=="Waterfall" & Date>="2018-05-16" & Date<="2018-06-11")

avgtempbypond<- avgtempearlyinstartopupae%>% group_by(Pond)  %>% summarise(meantemp=mean(meantemp, na.rm=TRUE))

#4
abioticsummarybypondwithsd<-allpondmetricsbysamplingstation %>% group_by(Pond) %>% dplyr::summarise(meanDepth=mean(Depth, na.rm=TRUE),sdDepth=sd(Depth, na.rm=TRUE), meanDOC=mean(DOC, na.rm=TRUE),sdDOC=sd(DOC, na.rm=TRUE), meanNH4=mean(NH4, na.rm=TRUE),sdNH4=sd(NH4, na.rm=TRUE), meanN03N02=mean(N03N02, na.rm=TRUE),sdN03N02=sd(N03N02, na.rm=TRUE), meanlogTotalP=mean(logTotalP, na.rm=TRUE),sdTotalP=sd(TotalP, na.rm=TRUE),meanlogConduc=mean(logConduc, na.rm=TRUE),sdConduc=sd(Conduc, na.rm=TRUE), meanDoPerc=mean(DoPerc, na.rm=TRUE),sdDoPerc=sd(DoPerc, na.rm=TRUE), meanpH=mean(pH, na.rm=TRUE),sdpH=sd(pH, na.rm=TRUE), meanFPOM_liter=mean(FPOM_liter, na.rm=TRUE),sdFPOM_liter=sd(FPOM_liter, na.rm=TRUE))

abioticsummarybypond<-allpondmetricsbysamplingstation %>% group_by(Pond) %>% dplyr::summarise(meanDepth=mean(Depth, na.rm=TRUE), meanDOC=mean(DOC, na.rm=TRUE), meanNH4=mean(NH4, na.rm=TRUE), meanN03N02=mean(N03N02, na.rm=TRUE), meanlogTotalP=mean(logTotalP, na.rm=TRUE),meanlogConduc=mean(logConduc, na.rm=TRUE), meanDoPerc=mean(DoPerc, na.rm=TRUE), meanpH=mean(pH, na.rm=TRUE),meanFPOM_liter=mean(FPOM_liter, na.rm=TRUE))

#5
abioticsummarybypond$Pond<-plyr::revalue(abioticsummarybypond$Pond, c("Vulgaris small"="Vulgaris.small"))
abioticandperimarea<- abioticsummarybypond %>% left_join(perimandareawide2, by=c("Pond"))
allabiotic<-abioticandperimarea %>% left_join(avgtempbypond, by=c("Pond"))

#6
allabiotic2 <- allabiotic[ -c(11:14,17)]
allabiotic2<-allabiotic2[,c(1, 11:13,2,14,10,3:6,9,7,8)] #change order of columns#

#7
subsetpredicators<- allabiotic2[-1]
correllationspondlevel<-correllations<-round(cor(subsetpredicators, use = "pairwise.complete.obs"),2)

```


```
