---
title: "Biofilm productivity"
author: "Melissa DeSiervo"
date: "5/6/2019"
output: html_document
---
This is the R code for developing linear modesl to predict Biofilm productivity from 2018 Greenland dataset. 

There are four datasets that go into this code. 2 are raw data files. 1 derived data file comes from the pond metrics.RMD, the last derived data file comes from the Biofilm metrics.RMD

1) Biofilm productivity == Derived dataset from the Biofilm metrics.RMD that calculates biofilm productivity at each sampling station
2) allpondmetricsbysamplingstation == raw data file (YSI data X 8 ponds X 3 sampling stations X 3 dates)
3) perimeterearea == raw data file perimeter and area data (8 ponds X 2 dates: pond thaw and after emergence)
4) temp and thermal sums = derived data file from Pond metrics.RMD that has the avg temps and thermal sums for each pond by date


First, upload the data
```{r Upload data}

#1

BiofilmproductivitySS<-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/4_Outputs/Biofilmmetrics_samplingstation.csv?token=AL4COXIEZYBXCZEGBORC7FK634DA4"))

BiofilmproductivitySS-as.data.frame(BiofilmproductivitySS)

#2

allpondmetricsbysamplingstation<-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/2_Inputs/allpondmetricsbysamplingstation.csv?token=AL4COXI5XA4YUHOLW6KPFG2634COY"))

allpondmetricsbysamplingstation-as.data.frame(allpondmetricsbysamplingstation)

#3
perimeterarea<-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/2_Inputs/perimeterandareaponds_2018.csv?token=AL4COXMXRF5HEGMXNDYHPEC634C4Y"))


perimeterarea<-as.data.frame(perimeterarea)
perimeterarea$Pond <- as.factor(perimeterarea$Pond)
perimeterarea$Date <- as.Date(perimeterarea$Date, "%m/%d/%Y")

#4

tempandthermalsums<-read.csv(url("https://raw.githubusercontent.com/melissadesiervo1031/Arctic_mosquito/master/4_Outputs/watertempthermalsums.csv?token=AL4COXLYFSOPERB77IOGIVS634CVY"))

tempandthermalsums<-as.data.frame(tempandthermalsums)
tempandthermalsums$Date <- as.Date(tempandthermalsums$Date, "%Y-%m-%d")


```

Install necessary packages
```{r Upload packages}
library(dplyr)
library(splitstackshape)
library(reshape2)
library(tidyr)
library(lme4)
library(MuMIn)
```


Pond metrics
1)subset to get just the last two measurements
2) average across the last two

```{r pond metrics}

allpondmetricslasttwo<-subset(allpondmetricsbysamplingstation, Instar=="2nd3rd"|Instar=="Pupae")

#2
pondmetricsSS<-allpondmetricslasttwo %>% group_by(Pond,SS2) %>% dplyr::summarise(meanDepth=mean(Depth, na.rm=TRUE),meanDOC=mean(DOC, na.rm=TRUE), meanNH4=mean(NH4, na.rm=TRUE), meanN03N02=mean(N03N02, na.rm=TRUE), meanlogTotalP=mean(logTotalP, na.rm=TRUE),meanlogConduc=mean(Conduc, na.rm=TRUE), meanDoPerc=mean(DoPerc, na.rm=TRUE), meanpH=mean(pH, na.rm=TRUE), meanFPOM_liter=mean(FPOM_liter, na.rm=TRUE))

pondmetricsSS2<<-pondmetricsSS %>% rename(Sampling.station = SS2)

pondmetricsSS<-as.data.frame(pondmetricsSS2)

```


Perimeter and area data
1)spread to wide format perimeter and area 
2)Calculate average perimeter, area, and loss
3) repeat each row three times to match other dataframes by SS (use expand rows from splitstackshape)
```{r perim and area data}
#1
perimeterwide<-perimeterarea %>% select(Pond,Measurement,Perimeter) %>% spread(Measurement,Perimeter) %>% dplyr::rename(Perimeter_1=Beginning, Perimeter_2=End)
areawide<-perimeterarea %>% select(Pond,Measurement,Area) %>% spread(Measurement,Area)%>% dplyr::rename(Area_1=Beginning, Area_2=End)

perimandareawide<- perimeterwide %>% left_join(areawide, by=c("Pond"))

#2
perimandareawide2<-perimandareawide %>% mutate(Perim_avg=(Perimeter_1 + Perimeter_2)/2, Area_avg=(Area_1+Area_2)/2, Perim_loss=abs(Perimeter_2-Perimeter_1), Area_loss=abs(Area_2-Area_1),  Area_loss_perc=(Area_loss/Area_1)*100)

#3
list3 <- rep("3",length(8)) ##make a vector of 3s (number of times to repeat##)
perimandareawide3 <- cbind(perimandareawide2, as.numeric(list3))
perimandareaSS<-expandRows(perimandareawide3, count=c(11))

```


Temp and thermal sums data
1) pull out the dates between the 2nd and third sampling date\
2) By pond: average temperature, pull out max thermal sum 0
3) repeat each row three times to match other dataframes by SS (use expand rows from splitstackshape)
```{r temp thermal sums}
#1
betweensecondandthird <- tempandthermalsums %>% dplyr::filter(Pond=="East" & Date>="2018-06-01" & Date<="2018-06-17"|
  Pond=="Golf" & Date>="2018-06-02" & Date<="2018-06-11"|Pond=="Ice" & Date>="2018-06-04" & Date<="2018-06-12"|
  Pond=="NoOil" & Date>="2018-05-27" & Date<="2018-06-04"|Pond=="Oil" & Date>="2018-05-13" & Date<="2018-06-05"|
  Pond=="Vulgaris" & Date>="2018-05-29" & Date<="2018-06-13"|Pond=="Vulgaris.small" & Date>="2018-06-04" & Date<="2018-06-13"|
  Pond=="Waterfall" & Date>="2018-05-27" & Date<="2018-06-11")

#2
tempthermalsums23<-betweensecondandthird %>% group_by(Pond) %>% dplyr::summarise(temp23=mean(meantemp,na.rm=TRUE), maxTS0=max(TS0, na.rm=TRUE))

#3
list3 <- rep("3",length(8)) ##make a vector of 3s (number of times to repeat##)
tempthermalsums23 <- cbind(tempthermalsums23, as.numeric(list3))
tempthermalsums23SS<-expandRows(tempthermalsums23, count=c(4))
tempthermalsums23SS<-as.data.frame(tempthermalsums23SS)
```

1) SS data merge with temperature data
2) merge #1 with perimeter and area data
3) merge #2 with the dependent predictors
```{r merge independent and dependent predictors}
#1
tempthermalsums23SS$Pond = droplevels(tempthermalsums23SS$Pond)
tempthermalsums23SS$Pond <-as.factor(tempthermalsums23SS$Pond)
tempthermalsums23SS<-cbind(tempthermalsums23SS, pondmetricsSS$Sampling.station)
colnames(tempthermalsums23SS)[4] <- 'Sampling.station'
pondmetricstempSS <- merge(pondmetricsSS,tempthermalsums23SS, by=c("Pond","Sampling.station"), sort=FALSE)
pondmetricstempSS$Pond<-plyr::revalue(pondmetricstempSS$Pond, c("Vulgaris small"="Vulgaris.small"))

#2
perimandareaSS2<-cbind(perimandareaSS, pondmetricsSS$Sampling.station)
colnames(perimandareaSS2)[11] <- c("Sampling.station")
pondmetricstempareaSS <- merge(pondmetricstempSS,perimandareaSS2,by=c("Pond","Sampling.station"), sort=FALSE)

#3
BiofilmproductivitySS2<-BiofilmproductivitySS%>%dplyr::rename(Sampling.station=Sampling.Station)

BiofilmproductivitySS2$Pond<-plyr::revalue(BiofilmproductivitySS2$Pond, c("Vulgaris small"="Vulgaris.small"))


allvariablesSS1 <- merge(pondmetricstempareaSS,BiofilmproductivitySS2,by=c("Pond","Sampling.station"), sort=FALSE)

##average by pond##

allvariablesbypond<-allvariablesSS1%>% group_by(Pond)%>% select_if(is.numeric)%>% summarise_each(funs(mean(., na.rm = TRUE)))

```


Make more models with 1 to 4 predictors. Then take the best models (AIC <2) and make an averaged model
1) make a correllation matrix and true/false for ones that are >0.5 correllation
2) add "na.omin" to the two full models (scaled and unscaled)
3) dredge function to make all combinations of models for unscaled data, subset the modesl <2
4) dredge function on the scaled data subset the modesl <2
5) Make the averaged model. == Table S4(?) in the Manuscript

Make models to predict biofilm productivity 
```{r make biofilm productivity models}


##subset out just the predictors#
subsetpredicators<- subset(allvariablesbypond, select =c(Perim_avg,Area_avg,Area_loss_perc,meanDepth,meanDOC,meanNH4,
meanN03N02,meanlogTotalP,meanlogConduc,meanDoPerc,meanpH,temp23,maxTS0,meanFPOM_liter))

#scale the predictors#
scaledpredictors<-scale(subsetpredicators)

#merge them back with the dependent variable in a dataframe#
biofilmdepnames<-subset(allvariablesbypond, select=c("Pond", "Prod.ug.cm.day"))
allvariablesscaled<-cbind(biofilmdepnames, scaledpredictors)

#1
smat3 <- abs(cor(subsetpredicators)) <= .5 #makes a correllation matrix and to say TRUE every time a value is greater than 0.5##
smat3[!lower.tri(smat3)] <- NA ##replaces the duplications with NA#

#2
fullunscaled<-lm(Prod.ug.cm.day~Perim_avg+Area_avg+Area_loss_perc+meanDepth+meanFPOM_liter+meanDOC+meanNH4+meanN03N02+meanlogTotalP+meanlogConduc+meanDoPerc+meanpH+meanpH+temp23+maxTS0, na.action = "na.fail", data=allvariablesbypond)

fullscaled<-lm(Prod.ug.cm.day~Perim_avg+Area_avg+Area_loss_perc+meanDepth+meanFPOM_liter+meanDOC+meanNH4+meanN03N02+meanlogTotalP+meanlogConduc+meanDoPerc+meanpH+meanpH+temp23+maxTS0, na.action = "na.fail", data=allvariablesscaled)

#3
dredgedmodelsallupto4noscale3 <- dredge(fullunscaled, subset=smat3, m.lim=c(1,4))
bestmodelsetnoscale3<-subset(dredgedmodelsallupto4noscale3, delta <2) 

#4
dredgedmodelsallupto4scale3 <- dredge(fullscaled, subset=smat3, m.lim=c(1,4))


###best model for predicting biofilm productivity is with perimeter##s

```

