---
title: "Biofilm productivity"
author: "Melissa DeSiervo"
date: "5/6/2019"
output: html_document
---
This is the R code for developing linear modesl to predict Biofilm productivity from 2018 Greenland dataset. 

There are four datasets that go into this code. 2 are raw data files. 1 derived data file comes from the pond metrics.RMD, the last derived data file comes from the Biofilm metrics.RMD

1) Biofilm productivity == Derived dataset from the Biofilm metrics.RMD that calculates biofilm productivity at each sampling station
2) allpondmetricsbysamplingstation == raw data file (YSI data X 8 ponds X 3 sampling stations X 3 dates)
3) perimeterearea == raw data file perimeter and area data (8 ponds X 2 dates: pond thaw and after emergence)
4) temp and thermal sums = derived data file from Pond metrics.RMD that has the avg temps and thermal sums for each pond by date


First, upload the data
```{r Upload data}

#1
BiofilmproductivitySS <- read.csv("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/Biofilmmetrics_samplingstation.csv")

#2
allpondmetricsbysamplingstation <- read.csv("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Inputs/allpondmetricsbysamplingstation.csv")
allpondmetricsbysamplingstation$Date <- as.Date(allpondmetricsbysamplingstation$Date, "%m/%d/%Y")

#3
perimeterarea <- read.csv("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Inputs/perimeterandareaponds_2018.csv",header = TRUE)
perimeterarea$Pond <- as.factor(perimeterarea$Pond)
perimeterarea$Date <- as.Date(perimeterarea$Date, "%m/%d/%Y")

#4
tempandthermalsums <- read.csv("C:/Users/Mitzi/Dropbox/Mosquito Science (1)/Data 2011 to 2018/Github repository/Arctic_mosquito/C-R paper/Data and analysis/Outputs/watertempthermalsums.csv")
tempandthermalsums$Date <- as.Date(tempandthermalsums$Date, "%Y-%m-%d")


```

Install necessary packages
```{r Upload packages}
library(dplyr)
library(splitstackshape)
library(lme4)
library(MuMIn)
```


Pond metrics
1)subset to get just the last two measurements
2) average across the last two

```{r pond metrics}

allpondmetricslasttwo<-subset(allpondmetricsbysamplingstation, Instar=="2nd3rd"|Instar=="Pupae")

#2
pondmetricsSS<-allpondmetricslasttwo %>% group_by(Pond, Sampling.station) %>% dplyr::summarise(meanDepth=mean(Depth, na.rm=TRUE),meanDOC=mean(DOC, na.rm=TRUE), meanNH4=mean(NH4, na.rm=TRUE), meanN03N02=mean(N03N02, na.rm=TRUE), meanlogTotalP=mean(logTotalP, na.rm=TRUE),meanlogConduc=mean(Conduc, na.rm=TRUE), meanDoPerc=mean(DoPerc, na.rm=TRUE), meanpH=mean(pH, na.rm=TRUE), meanFPOM_liter=mean(FPOM_liter, na.rm=TRUE))

pondmetricsSS<-as.data.frame(pondmetricsSS)

```


Perimeter and area data
1)spread to wide format perimeter and area 
2)Calculate average perimeter, area, and loss
3) repeat each row three times to match other dataframes by SS (use expand rows from splitstackshape)
```{r perim and area data}
#1
perimeterwide<-perimeterarea %>% select(Pond,Measurement,Perimeter) %>% spread(Measurement,Perimeter)%>% rename(Perimeter_1=Beginning, Perimeter_2=End)

areawide<-perimeterarea %>% select(Pond,Measurement,Area) %>% spread(Measurement,Area)%>% rename(Area_1=Beginning, Area_2=End)

perimandareawide<- perimeterwide %>% left_join(areawide, by=c("Pond"))

#2
perimandareawide2<-perimandareawide %>% mutate(Perim_avg=(Perimeter_1 + Perimeter_2)/2, Area_avg=(Area_1+Area_2)/2, Perim_loss=abs(Perimeter_2-Perimeter_1), Area_loss=abs(Area_2-Area_1),  Area_loss_perc=(Area_loss/Area_1)*100)

#3
list3 <- rep("3",length(8)) ##make a vector of 3s (number of times to repeat##)
perimandareawide3 <- cbind(perimandareawide2, as.numeric(list3))
perimandareaSS<-expandRows(perimandareawide3, count=c(11))

```


Temp and thermal sums data
1) pull out the dates between the 2nd and third sampling date\
2) By pond: average temperature, pull out max thermal sum 0
3) repeat each row three times to match other dataframes by SS (use expand rows from splitstackshape)
```{r temp thermal sums}
#1
betweensecondandthird <- tempandthermalsums %>% filter(Pond=="East" & Date>="2018-06-01" & Date<="2018-06-17"|
  Pond=="Golf" & Date>="2018-06-02" & Date<="2018-06-11"|Pond=="Ice" & Date>="2018-06-04" & Date<="2018-06-12"|
  Pond=="NoOil" & Date>="2018-05-27" & Date<="2018-06-04"|Pond=="Oil" & Date>="2018-05-13" & Date<="2018-06-05"|
  Pond=="Vulgaris" & Date>="2018-05-29" & Date<="2018-06-13"|Pond=="Vulgaris.small" & Date>="2018-06-04" & Date<="2018-06-13"|
  Pond=="Waterfall" & Date>="2018-05-27" & Date<="2018-06-11")

#2
tempthermalsums23<-betweensecondandthird %>% group_by(Pond) %>% dplyr::summarise(temp23=mean(meantemp,na.rm=TRUE), maxTS0=max(TS0, na.rm=TRUE))

#3
list3 <- rep("3",length(8)) ##make a vector of 3s (number of times to repeat##)
tempthermalsums23 <- cbind(tempthermalsums23, as.numeric(list3))
tempthermalsums23SS<-expandRows(tempthermalsums23, count=c(4))
tempthermalsums23SS<-as.data.frame(tempthermalsums23SS)
```

1) SS data merge with temperature data
2) merge #1 with perimeter and area data
3) merge #2 with the dependent predictors
```{r merge independent and dependent predictors}
#1
tempthermalsums23SS$Pond = droplevels(tempthermalsums23SS$Pond)
tempthermalsums23SS$Pond <-as.factor(tempthermalsums23SS$Pond)
tempthermalsums23SS<-cbind(tempthermalsums23SS, pondmetricsSS$Sampling.station)
colnames(tempthermalsums23SS)[4] <- 'Sampling.station'
pondmetricstempSS <- merge(pondmetricsSS,tempthermalsums23SS, by=c("Pond","Sampling.station"), sort=FALSE)

#2
perimandareaSS<-cbind(perimandareaSS, pondmetricsSS$Sampling.station)
colnames(perimandareaSS)[11] <- 'Sampling.station'

pondmetricstempareaSS <- merge(pondmetricstempSS,perimandareaSS,by=c("Pond","Sampling.station"), sort=FALSE)

#3
pondmetricstempareaSS$Pond<-plyr::revalue(pondmetricstempareaSS$Pond, c("Vulgaris.small"="Vulgaris small"))
BiofilmproductivitySS<-BiofilmproductivitySS  %>% rename(Sampling.station=Sampling.Station)
allvariablesSS1 <- merge(pondmetricstempareaSS,BiofilmproductivitySS,by=c("Pond","Sampling.station"), sort=FALSE)

```


Make a bunch of models to predict biofilm productivity 
```{r make biofilm productivity models}

allvariablesSS<-allvariablesSS1[!is.na(allvariablesSS1$Prod.ug.cm.day), ]


#null#
Nullmod<-lmer(Prod.ug.cm.day~ (1|Pond), REML=F, data=allvariablesSS)

##univariate models## ##nmote I had to to REM=f to be able to rank with AIC#
Perim<-lmer(Prod.ug.cm.day~Perim_avg + (1|Pond), REML=F, data=allvariablesSS)
Area<-lmer(Prod.ug.cm.day~ Area_avg+(1|Pond), REML=F, data=allvariablesSS)
Area_loss<-lmer(Prod.ug.cm.day~Area_loss_perc+(1|Pond), REML=F, data=allvariablesSS)
Depth<-lmer(Prod.ug.cm.day~meanDepth+(1|Pond), REML=F, data=allvariablesSS)

FPOM<-lmer(Prod.ug.cm.day~meanFPOM_liter+(1|Pond), REML=F, data=allvariablesSS)
DOC<-lmer(Prod.ug.cm.day~meanDOC+(1|Pond), REML=F, data=allvariablesSS)
NH4<-lmer(Prod.ug.cm.day~meanNH4+(1|Pond), REML=F, data=allvariablesSS)
N03N02<-lmer(Prod.ug.cm.day~meanN03N02+(1|Pond), REML=F, data=allvariablesSS)
totalP<-lmer(Prod.ug.cm.day~meanlogTotalP+(1|Pond), REML=F, data=allvariablesSS)

Conduct<-lmer(Prod.ug.cm.day~meanlogConduc+(1|Pond), REML=F, data=allvariablesSS)
DOperc<-lmer(Prod.ug.cm.day~meanDoPerc+(1|Pond), REML=F, data=allvariablesSS)
pH<-lmer(Prod.ug.cm.day~meanpH+(1|Pond), REML=F, data=allvariablesSS)
temphobo<-lmer(Prod.ug.cm.day~temp23+(1|Pond), REML=F, data=allvariablesSS)
maxTS0<-lmer(Prod.ug.cm.day~maxTS0+(1|Pond), REML=F, data=allvariablesSS)

#Full##
fullunscaled<-lmer(Prod.ug.cm.day~Perim_avg+Area_avg+Area_loss_perc+meanDepth+meanFPOM_liter+meanDOC+meanNH4+meanN03N02+meanlogTotalP+meanlogConduc+meanDoPerc+meanpH+meanpH+temp23+maxTS0+(1|Pond), REML=F, data=allvariablesSS)

```


Scale data and repeate process Make a bunch of models to predict biofilm productivity 
```{r make biofilm productivity models-scaled data}

##subset out just the predictors#
subsetpredicators<- subset(allvariablesSS1, select =c(Perim_avg,Area_avg,Area_loss_perc,meanDepth,meanDOC,meanNH4,
meanN03N02,meanlogTotalP,meanlogConduc,meanDoPerc,meanpH,temp23,maxTS0,meanFPOM_liter))

#scale the predictors#
scaledpredictors<-scale(subsetpredicators)

#merge them back with the dependent variable in a dataframe#
biofilmdepnames<-subset(BiofilmproductivitySS, select=c("Pond", "Prod.ug.cm.day"))
allvariablesSSscaled<-cbind(biofilmdepnames, scaledpredictors)
allvariablesSSscaled<-allvariablesSS[!is.na(allvariablesSS$Prod.ug.cm.day), ]

#null#
Nullmod2<-lmer(Prod.ug.cm.day~ (1|Pond), REML=F, data=allvariablesSSscaled)

##univariate models## ##nmote I had to to REM=f to be able to rank with AIC#
Perim2<-lmer(Prod.ug.cm.day~Perim_avg + (1|Pond), REML=F, data=allvariablesSSscaled)
Area2<-lmer(Prod.ug.cm.day~ Area_avg+(1|Pond), REML=F, data=allvariablesSSscaled)
Area_loss2<-lmer(Prod.ug.cm.day~Area_loss_perc+(1|Pond), REML=F, data=allvariablesSSscaled)
Depth2<-lmer(Prod.ug.cm.day~meanDepth+(1|Pond), REML=F, data=allvariablesSSscaled)

FPOM2<-lmer(Prod.ug.cm.day~meanFPOM_liter+(1|Pond), REML=F, data=allvariablesSSscaled)
DOC2<-lmer(Prod.ug.cm.day~meanDOC+(1|Pond), REML=F, data=allvariablesSSscaled)
NH42<-lmer(Prod.ug.cm.day~meanNH4+(1|Pond), REML=F, data=allvariablesSSscaled)
N03N022<-lmer(Prod.ug.cm.day~meanN03N02+(1|Pond), REML=F, data=allvariablesSSscaled)
totalP2<-lmer(Prod.ug.cm.day~meanlogTotalP+(1|Pond), REML=F, data=allvariablesSSscaled)

Conduct2<-lmer(Prod.ug.cm.day~meanlogConduc+(1|Pond), REML=F, data=allvariablesSSscaled)
DOperc2<-lmer(Prod.ug.cm.day~meanDoPerc+(1|Pond), REML=F, data=allvariablesSSscaled)
pH2<-lmer(Prod.ug.cm.day~meanpH+(1|Pond), REML=F, data=allvariablesSSscaled)
temphobo2<-lmer(Prod.ug.cm.day~temp23+(1|Pond), REML=F, data=allvariablesSSscaled)
maxTS02<-lmer(Prod.ug.cm.day~maxTS0+(1|Pond), REML=F, data=allvariablesSSscaled)

#Full##
fullscaled<-lmer(Prod.ug.cm.day~Perim_avg+Area_avg+Area_loss_perc+meanDepth+meanFPOM_liter+meanDOC+meanNH4+meanN03N02+meanlogTotalP+meanlogConduc+meanDoPerc+meanpH+meanpH+temp23+maxTS0+(1|Pond), REML=F, data=allvariablesSSscaled)

```

AIC table function
```{r AIC table}
aictable<-function(X,m){
  #
  #  This function will create a full model selection table based on AICc.
  #  
  #  Inputs:
  #
  #  X is the AIC table produced by R by using AIC(model1,model2, ...)
  #  m is the sample size
  #
  #
  rnames<-row.names(X)
  AICc<-X$AIC+2*X$df*(X$df+1)/(m-X$df-1)     #small-sample correction
  logL<-X$df-X$AIC/2                         #Log-likelihood
  tab<-data.frame(X[,1],logL,AICc)           #Remove AIC column; add logL and AICc
  colnames(tab)[1]<-c("Params")              #Rename "df" column   
  row.names(tab)<-rnames
  tab<-tab[order(tab$AICc),]                 #Sort by ascending AICc value
  deltaAICc<-tab$AICc-min(tab$AICc)          #Delta AICc
  weight<-exp(-deltaAICc/2)/sum(exp(-deltaAICc/2))  #Weights
  cumwt<-weight                              #Column for cumulative weight
  for(i in 2:dim(X)[1]){                  
    cumwt[i]<-cumwt[i-1]+cumwt[i]              #Accumulate weight from the top
  }
  tab<-data.frame(tab,deltaAICc,weight,cumwt)
  tab<-round(tab,4)
  tab
}

```

AIC tables for univariate models, null, full for scaled and unscaled data 
This is NOT reported in manuscript
```{r AIC table for scaled and unscaled data}
rawaicunscaled<-AIC(Nullmod,Perim,Area,Area_loss,Depth,DOC,NH4,N03N02,totalP,DOperc, Conduct,pH,temphobo, maxTS0, FPOM,fullunscaled)
nR<-dim(allvariablesSS)[1]  #Sample size 
aictableunscaled<-aictable(rawaicunscaled,nR)

rawaicscaled<-AIC(Nullmod2,Perim2,Area2,Area_loss2,Depth2,DOC2,NH42,N03N022,totalP2,DOperc2, Conduct2,pH2,temphobo2, maxTS02, FPOM2,fullscaled)
nR<-dim(allvariablesSSscaled)[1]  #Sample size 
aictablescaled<-aictable(rawaicscaled,nR)

```

Make more models with 1 to 4 predictors. Then take the best models (AIC <2) and make an averaged model
1) make a correllation matrix and true/false for ones that are >0.5 correllation
2) add "na.omin" to the two full models (scaled and unscaled)
3) dredge function to make all combinations of models for unscaled data, subset the modesl <2
4) dredge function on the scaled data subset the modesl <2
5) Make the averaged model. == Table S4(?) in the Manuscript

```{r make more models and average}
#1
smat3 <- abs(cor(subsetpredicators)) <= .5 #makes a correllation matrix and to say TRUE every time a value is greater than 0.5##
smat3[!lower.tri(smat3)] <- NA ##replaces the duplications with NA#

#2
fullunscaled<-lmer(Prod.ug.cm.day~Perim_avg+Area_avg+Area_loss_perc+meanDepth+meanFPOM_liter+meanDOC+meanNH4+meanN03N02+meanlogTotalP+meanlogConduc+meanDoPerc+meanpH+meanpH+temp23+maxTS0+(1|Pond), REML=F,na.action = "na.fail", data=allvariablesSS)

fullscaled<-lmer(Prod.ug.cm.day~Perim_avg+Area_avg+Area_loss_perc+meanDepth+meanFPOM_liter+meanDOC+meanNH4+meanN03N02+meanlogTotalP+meanlogConduc+meanDoPerc+meanpH+meanpH+temp23+maxTS0+(1|Pond), REML=F,na.action = "na.fail", data=allvariablesSSscaled)

#3
dredgedmodelsallupto4noscale3 <- dredge(fullunscaled, subset=smat3, m.lim=c(1,4))
bestmodelsetnoscale3<-subset(dredgedmodelsallupto4noscale3, delta <2) 

#4
dredgedmodelsallupto4scale3 <- dredge(fullscaled, subset=smat3, m.lim=c(1,4))
bestmodelsetscale3<-subset(dredgedmodelsallupto4scale3, delta <2) 

#5
avgmodelnoscale3<-model.avg(bestmodelsetnoscale3)
avgmodelscale3<-model.avg(bestmodelsetscale3)

```

Calculate the adjusted Ys and make Figure
```{r Adjusted ys and figure}

```




